---
title: "Microwave_Analysis_Microglia"
output: html_document
date: "2024-03-14"
---


This script is designed to put together csv files from CellProfiler pipline which measures morphological parameters of individual cells

STEP-1  Microscopic image aqusition 
Images are acquired post microwave (MW) exposure for stimulation condition and without exposure for sham condition using EVOS M7000 microscope
respective channel ID:
GFP - lentiviral infected amCyan Microglia
DAPI - Nuclei stainning

STEP-2  CellProfiler pipeline (V-4.2.6) (MacOS)
Example of pipeline is provided on github repo


STEP-3  Downstream data analysis using the .csv files from CellProfiler output module





```{r}
############################################################################################################################################
#                                                                                                                                          #
#                                                          Package Validation                                                              #
#                                                                                                                                          #
############################################################################################################################################

# List to store package names
packages <- list()

# Specify the required packages
packages$my_packages <- c("readr", "plyr", "readxl", "dplyr", "factoextra", "cluster", "tidyverse", "corrplot", "tidyr", "ggplot2", "Biobase", "BiocManager", "pheatmap", "vroom", "Seurat", "recipes", "tidytext", "embed", "corrr", "viridis", "randomForest", "BiocParallel", "pheatmap", "dendextend", "RColorBrewer", "dendsort", "ggExtra","remotes", "ggstream","smplot2", "devtools", "ggthemes", "Seurat", "plotly", "shiny", "seriation", "ggdist", "reshape2")                                       

# Check which packages are not installed
packages$not_installed <- packages$my_packages[!(packages$my_packages %in% installed.packages()[ , "Package"])]

# Install the packages that are not already installed
if(length(packages$not_installed)) install.packages(packages$not_installed)   

##### BIOCONDUCTOR BASED PACKAGES #####

# Check if BiocManager is available, if not, install it
if (!require("BiocManager", quietly = TRUE))
  install.packages("BiocManager")

# Install specific Bioconductor packages
BiocManager::install(version = "3.18")
BiocManager::install("ConsensusClusterPlus")
BiocManager::install("BiocParallel")

##### LOAD ALL THE PACKAGES AT ONCE #####

# Load all the required packages
lapply(packages$my_packages, require, character.only = TRUE)

```

```{r}


############################################################################################################################################
#                                                                                                                                          #
#                                                          Data Import                                                                     #
#                                                                                                                                          #
############################################################################################################################################

# Initialize the 'import' list to store data frames
import <- list()

# Importing data from a CSV file and removing unnecessary columns
import$df_all = read_csv("~/Desktop/DIAQNOS/VJ_Microwave_EXP/Microglia/Analysis/Datasheet/df_all_combined")

import$df_all <- subset(import$df_all, select = -c(Center_Z_soma, Center_Z_cell, PathName_OG_cell, BoundingBoxMaximum_X_cell, BoundingBoxMaximum_Y_cell,
                                                  BoundingBoxMinimum_X_cell, BoundingBoxMinimum_Y_cell, Number_Object_Number_cell, Parent_IdentifyPrimaryObjects_cell,
                                                  PathName_OG_soma, Children_IdentifySecondaryObjects_Count_soma, Center_X_soma.1, Center_Y_soma.1, Number_Object_Number_soma,
                                                  BoundingBoxMaximum_X_soma, BoundingBoxMaximum_Y_soma, BoundingBoxMinimum_X_soma, BoundingBoxMinimum_Y_soma,
                                                  MaxX_OrigBlue_soma, MaxY_OrigBlue_soma, MaxZ_OrigBlue_soma, EulerNumber_cell, Extent_cell, MaximumRadius_cell,
                                                  Orientation_cell, IntegratedIntensityEdge_OrigGreen_cell, LowerQuartileOrigGreen_cell, MADOrigGreen_cell,
                                                  MassDisplacement_OrigGreen_cell, MaxIntensityEdge_OrigGreen_cell, MedianOrigGreen_cell, MinIntensityEdge_OrigGreen_cell, StdIntensityEdge_OrigGreen_cell,
                                                  StdOrigGreen_cell, UpperQuartileOrigGreen_cell, EulerNumber_soma, Extent_soma, MaximumRadius_soma, Orientation_soma,
                                                  IntegratedIntensityEdge_OrigBlue_soma, LowerQuartileOrigBlue_soma, MADOrigBlue_soma, MassDisplacement_OrigBlue_soma,
                                                  MaxIntensityEdge_OrigBlue_soma, MedianOrigBlue_soma, MinIntensityEdge_OrigBlue_soma, StdIntensityEdge_OrigBlue_soma, StdOrigBlue_soma,
                                                  UpperQuartileOrigBlue_soma, CenterMassX_OrigBlue_soma, CenterMassY_OrigBlue_soma, CenterMassZ_OrigBlue_soma, Center_X_cell...12,
                                                  Center_Y_cell...13, Center_Y_cell...47, Center_X_cell...46 ))

import$df_all <- import$df_all %>%
  # Extracting file name components into separate columns
  mutate(FileName_OG_cell = str_remove(FileName_OG_cell, pattern = ".tif"))

import$colmn <- paste(c("Cell_Type","MW", "Expo_Time", "Recovery_Time", "Plate_Number", "Plate_Name", "Image_Area"))

import$df_all <- tidyr::separate(
  data = import$df_all,
  col = FileName_OG_cell,
  sep = "_",
  into = import$colmn,
  remove = FALSE)

import$df_all$Condition <- substr(import$df_all$Plate_Name, 1, 2)

# Renaming columns for ease of use
names(import$df_all)[names(import$df_all) == 'ObjectSkeleton_NumberBranchEnds_MorphologicalSkeleton_soma'] <- 'Branch_Ends'
names(import$df_all)[names(import$df_all) == 'ObjectSkeleton_NumberNonTrunkBranches_MorphologicalSkeleton_soma'] <- 'Non_Trunk_Branch'
names(import$df_all)[names(import$df_all) == 'ObjectSkeleton_NumberTrunks_MorphologicalSkeleton_soma'] <- 'Trunk_Branch'
names(import$df_all)[names(import$df_all) == 'ObjectSkeleton_TotalObjectSkeletonLength_MorphologicalSkeleton_soma'] <- 'Skeleton_Length'

# Calculating ramification index of the cell
import$df_all$RI <- ((import$df_all$Perimeter_cell / import$df_all$Area_cell) / (2*sqrt((pi / import$df_all$Area_cell))))

# Calculating area ratio of cell to soma
import$df_all$area_ratio <- import$df_all$Area_cell / import$df_all$Area_soma

# Calculating length to width ratio of cell & soma
import$df_all$Length_Width_Ratio_cell <- import$df_all$MaxFeretDiameter_cell / import$df_all$MinFeretDiameter_cell
import$df_all$Length_Width_Ratio_soma <- import$df_all$MaxFeretDiameter_soma / import$df_all$MinFeretDiameter_soma

# Calculating aspect ratio of cell and soma
import$df_all$Aspect_Ratio_cell <- import$df_all$MajorAxisLength_cell / import$df_all$MinorAxisLength_cell
import$df_all$Aspect_Ratio_soma <- import$df_all$MajorAxisLength_soma / import$df_all$MinorAxisLength_soma

# Calculating cytoplasmic area of microglia
import$df_all$Cyto_Area <- import$df_all$Area_cell - import$df_all$Area_soma

# Removing NA values from the dataframe
import$df_all <- na.omit(import$df_all)

# Encoding "MW" column with descriptive labels
import$df_all$MW <- case_when(import$df_all$MW =="C" ~ "Sham",
                             import$df_all$MW=="S" ~ "Stimulation")

# Categorizing "Region" column into "Periphery" and "Center" based on image area
import$df_all$Region <- case_when(import$df_all$Image_Area %in% c("1", "2", "3", "5", "6", "8", "9", "10") ~ "Periphery",
                                 import$df_all$Image_Area %in% c("4", "7") ~ "Center")

# Selecting relevant columns for further analysis
import$df_all <- import$df_all %>% dplyr::select(c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 33, 34, 53, 61), everything())

```


```{r}
############################################################################################################################################
#                                                                                                                                          #
#                                                Cell Proliferation analysis                                                               #
#                                                                                                                                          #
############################################################################################################################################


# Create a new object to store the results of counting
count <- list()

# Count the number of cells for each condition
count$df_counts <- import$df_all %>% 
  group_by(ImageNumber_cell, Plate_Name, Condition, Image_Area, Recovery_Time, Expo_Time, MW, Region) %>% 
  summarize(num_cells = n())

# Define the area of the image
count$df_counts$Image_area <- 3768 * 3088

# Create a new box plot showing the normalized cell count for different conditions
p1 <- ggplot(count$df_counts, aes(x = MW, y = 10000*(num_cells / Image_area), fill = MW)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.85) +
  geom_jitter(width = 0.1, aes(color = as.factor(Region))) +
  ggpubr::stat_compare_means(comparisons = list(c("Sham","Stimulation")), method = "t.test") +
  theme_bw() + 
  facet_grid(Expo_Time ~ Recovery_Time) +
  xlab("") +
  ylim(c(0, 6)) +
  ylab("Normalized cell count") +
  ggtitle("Cell Proliferation index overall")+
  scale_color_manual(values = c("blue", "green")) +
  scale_fill_manual(values = c("darkgrey", "red")) +
  labs(color = "Region") +
  theme(
    plot.title = element_text(size = 12, hjust = 0.5, face = "bold"),
    axis.title.x = element_text(size = 12, face = "bold"),
    axis.title.y = element_text(size = 12, face = "bold"),
    axis.text.x = element_text(size = 10, face = "bold", angle = 45, hjust = 1),
    axis.text.y = element_text(size = 10, face = "bold"),
    legend.text = element_text(size = 9, face = "bold"),
    legend.title = element_text(size = 9, face = "bold"),
    legend.key.size = unit(1.5, "lines"),
    legend.position = "right",
    strip.text = element_text(size = 10, face = "bold")
  )

p2 <- ggplot(count$df_counts, aes(x = MW, y = 10000*(num_cells / Image_area), fill = MW)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.85) +
  geom_jitter(width = 0.1) +
  ggpubr::stat_compare_means(comparisons = list(c("Sham","Stimulation")), method = "t.test") +
  theme_bw() + 
  facet_grid(Region ~ Recovery_Time) +
  xlab("") +
  ylim(c(0, 6)) +
  ylab("Normalized cell count") +
  ggtitle("Cell Proliferation index center vs periphery")+
  scale_fill_manual(values = c("darkgrey", "red")) +
  labs(color = "Region") +
  theme(
    plot.title = element_text(size = 12, hjust = 0.5, face = "bold"),
    axis.title.x = element_text(size = 12, face = "bold"),
    axis.title.y = element_text(size = 12, face = "bold"),
    axis.text.x = element_text(size = 10, face = "bold", angle = 45, hjust = 1),
    axis.text.y = element_text(size = 10, face = "bold"),
    legend.text = element_text(size = 9, face = "bold"),
    legend.title = element_text(size = 9, face = "bold"),
    legend.key.size = unit(1.5, "lines"),
    legend.position = "right",
    strip.text = element_text(size = 10, face = "bold")
  )

cowplot::plot_grid(p1, p2)



```


```{r}

############################################################################################################################################
#                                                                                                                                          #
#                                                        Parameter plotting.                                                               #
#                                                                                                                                          #
############################################################################################################################################

# Create a parameter plot showing the values of different para meters from your data set

# replace the name of the parameter on y axis 

ggplot(import$df_all, aes(x = MW, y = RI, fill = MW)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.75, width = 0.20) +
  stat_halfeye(
    adjust = 1,  # Adjust bandwidth
    justification = -0.05,  # Move to the right
    .width = 0,  # Remove the slub interval
    alpha = 0.4,
    point_colour = NA
  ) +
  theme_bw() + 
  facet_grid(Region ~ Recovery_Time) +
  xlab("") +
  scale_color_manual(values = c("blue", "green")) +
  scale_fill_manual(values = c("darkgrey", "red")) +
  labs(fill = "Microwave", color = "Exposure time (mins)") +
  ylim(c(0.2, 1.5)) +
  theme(
    plot.title = element_text(size = 12, hjust = 0.5, face = "bold"),
    axis.title.x = element_text(size = 12, face = "bold"),
    axis.title.y = element_text(size = 12, face = "bold"),
    axis.text.x = element_text(size = 10, face = "bold", angle = 45, hjust = 1),
    axis.text.y = element_text(size = 10, face = "bold"),
    legend.text = element_text(size = 9, face = "bold"),
    legend.title = element_text(size = 9, face = "bold"),
    legend.key.size = unit(1.5, "lines"),
    legend.position = "right",
    strip.text = element_text(size = 10, face = "bold")
  )


```

```{r}

############################################################################################################################################
#                                                                                                                                          #
#                                                   Hierarchy clustering analysis                                                          #
#                                                                                                                                          #
############################################################################################################################################



# Create a new object called H_clust
H_clust <- list()

# Assign the data frame from import to H_clust
H_clust$df_clust <- import$df_all

# Scale the columns
H_clust$scale <- scale(H_clust$df_clust[, 15:61])
H_clust$scaled_df <- cbind(H_clust$df_clust[, 1:14], H_clust$scale)

# Perform hierarchical clustering on the scaled data
H_clust$cluster_cols <- hclust(dist(t(H_clust$scaled_df[, 15:61])))

# Define a function to sort the hierarchical clustering result
H_clust$sort_hclust <- function(...) as.hclust(dendsort(as.dendrogram(...)))

# Sort the hierarchical clustering result
H_clust$cluster_cols <- H_clust$sort_hclust(H_clust$cluster_cols)

# Plot the sorted dendrogram
H_clust$gobal_dendrogram <- fviz_dend(H_clust$cluster_cols, cex = 0.8, k = 5, 
                                       rect = TRUE,  
                                       k_colors = "jco",
                                       rect_border = "jco", 
                                       rect_fill = TRUE, 
                                       horiz = TRUE) +
  theme(
    plot.title = element_text(size = 12, hjust = 0.5, face = "bold"),
    axis.title.x = element_text(size = 10, face = "bold"),
    axis.title.y = element_text(size = 10, face = "bold"),
    axis.text.x = element_text(size = 11, face = "bold"),
    axis.text.y  = element_text(size = 11, face = "bold"),
    legend.text = element_text(size = 8, face = "bold"),
    legend.title = element_text(size = 8, face = "bold"),
    legend.key.size = unit(1.5, "lines"),
    legend.position = "bottom",
    strip.text = element_text(size = 10, face = "bold"))

# Plot the global dendrogram
plot(H_clust$gobal_dendrogram)

```





```{r}

############################################################################################################################################
#                                                                                                                                          #
#                                                   K-means clustering analysis                                                            #
#                                                                                                                                          #
############################################################################################################################################


# Create a new object called PCA
PCA <- list()

# Assign the data frame from import to PCA
PCA$df_pca <- import$df_all

# Find kmeans based on the optimal number of clusters
set.seed(1000)
PCA$kmeans_all <- kmeans(PCA$df_pca[, 15:61], centers = 6, nstart = 25)
PCA$kmeans_all

# Add the cluster information to the data frame
PCA$df_pca$Cluster <- PCA$kmeans_all$cluster

PCA$df_pca <- PCA$df_pca[which(PCA$df_pca$Cluster == 6),]

# Find kmeans based on the optimal number of clusters
set.seed(1000)
PCA$kmeans_all <- kmeans(PCA$df_pca[, 15:61], centers = 6, nstart = 25)
PCA$kmeans_all

# Add the cluster information to the data frame
PCA$df_pca$Cluster <- PCA$kmeans_all$cluster

PCA$df_pca <- PCA$df_pca[which(PCA$df_pca$Cluster == 2),]

# Find kmeans based on the optimal number of clusters
set.seed(1000)
PCA$kmeans_all <- kmeans(PCA$df_pca[, 15:61], centers = 6, nstart = 25)
PCA$kmeans_all

# Add the cluster information to the data frame
PCA$df_pca$Cluster <- PCA$kmeans_all$cluster
```
```{r}


############################################################################################################################################
#                                                                                                                                          #
#                                                   PCA reduction & visualization                                                          #
#                                                                                                                                          #
############################################################################################################################################

## PCA starts here

PCA$pca_rec <- recipe(~., data = PCA$df_pca) %>%
  update_role(Center_X_soma, Center_Y_soma, Cell_Type, MW, Expo_Time, Recovery_Time, Plate_Number, Plate_Name, Image_Area, Condition,
               Cluster, FileName_OG_cell, ImageNumber_cell, ObjectNumber_cell, Region, new_role = "id") %>%
  step_normalize(all_predictors()) %>%
  step_pca(all_predictors())

PCA$pca_prep <- prep(PCA$pca_rec)

PCA$pca_prep


PCA$tidied_pca <- tidy(PCA$pca_prep, 2)

PCA$tidied_pca %>%
  filter(component %in% paste0("PC", 1:5)) %>%
  mutate(component = fct_inorder(component)) %>%
  ggplot(aes(value, terms, fill = terms)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~component, nrow = 1) +
  labs(y = NULL)


PCA$tidied_pca %>%
  filter(component %in% paste0("PC", 1:5)) %>%
  group_by(component) %>%
  top_n(15, abs(value)) %>%
  ungroup() %>%
  mutate(terms = reorder_within(terms, abs(value), component)) %>%
  ggplot(aes(abs(value), terms, fill = value > 0)) +
  geom_col() +
  facet_wrap(~component, scales = "free_y") +
  scale_y_reordered() +
  labs(
    x = "Absolute value of contribution",
    y = NULL, fill = "Positive?"
  )




juice(PCA$pca_prep) %>%
  ggplot(aes(PC1, PC2, label = NA)) +
  geom_point(aes(color = as.factor(Cluster)), alpha = 0.7, size = 2, shape = ".") +
  geom_text(check_overlap = TRUE, hjust = "inward", family = "IBMPlexSans") +
  labs(color = NULL)+
  facet_grid(~Cluster)+
  xlim(-30, 65)+
  ylim(-40, 60)+
  scale_color_manual(values=company_colors)+
  geom_density2d(geom="polygon")+
  theme_classic()+
  geom_vline(xintercept = 0, linetype = "dotted")+ # Vertical line
  geom_hline(yintercept = 0, linetype = "dotted")+ 
  ggtitle("Major morpho-families of microglia")+
  theme(
    plot.title = element_text(size=12, hjust = 0.5, face="bold"),
    axis.title.x = element_text(size=10, face="bold"),
    axis.title.y = element_text(size=10, face="bold"),
    axis.text.x = element_text(size = 8, face="bold"),
    axis.text.y  = element_text(size = 8, face="bold"),
    legend.text = element_text(size = 8,  face="bold"),
    legend.title = element_text(size = 10,  face="bold"),
    legend.key.size = unit(1.5, "lines"),
    legend.position = "right",
    strip.text = element_text(size = 10, face = "bold"))


```
```{r}

############################################################################################################################################
#                                                                                                                                          #
#                                                   UMAP reduction & visualization                                                         #
#                                                                                                                                          #
############################################################################################################################################


# Create a recipe for UMAP
umap_rec <- recipe(~ ., data = PCA$df_pca) %>%
  update_role(Center_X_soma, Center_Y_cell, Cell_Type, MW, Expo_Time, Recovery_Time, Plate_Number, Plate_Name, Image_Area, Condition,
              Cluster, FileName_OG_cell, ImageNumber_cell, ObjectNumber_cell, new_role, Region = "id") %>%
  step_normalize(all_predictors()) %>%
  step_umap(all_predictors())

# Prepare UMAP data
umap_prep <- prep(umap_rec)
umap_prep

# Plot UMAP
juice(umap_prep) %>%
  ggplot(aes(UMAP1, UMAP2, label = NA)) +
  geom_point(aes(color = as.factor(Cluster)), alpha = 0.7, size = 2) +
  geom_text(check_overlap = TRUE, hjust = "inward", family = "IBMPlexSans") +
  labs(color = NULL)




```



``````{r}

############################################################################################################################################
#                                                                                                                                          #
#                                                   Variable parameters within cluster                                                     #
#                                                                                                                                          #
############################################################################################################################################

##### IMPORTANCE PARAMETER PLOT #####

# Calculate the importance of each column in each condition
H_clust$importance <- aggregate(PCA$df_pca[, 15:62], by = list(Weeks = PCA$df_pca$Cluster), FUN = mean)

# Melt the data frame to long format
H_clust$importance_melted <- melt(H_clust$importance, id.vars = c("Weeks"), variable.name = "Parameter", value.name = "Importance")

# Group the melted data frame
H_clust$df_grouped <- H_clust$importance_melted %>% group_by(Weeks)

# Select the top 20 parameters
H_clust$f_top50 <- H_clust$df_grouped %>% 
  slice_max(order_by = Importance, n = 50) %>%
  ungroup()

# Create a bar plot for the top 20 parameters
H_clust$top50_parameter <- ggplot(H_clust$f_top50, aes(x = Importance, y = Parameter, fill = factor(Weeks))) + 
  geom_col() +
  facet_grid(~Weeks) +
  scale_fill_manual(values = company_colors) +
  labs(title = "Top 20 Parameters") +
  theme_bw() +
  labs(fill = "Cluster") +
  theme(
    plot.title = element_text(size = 14, hjust = 0.5, face = "bold"),
    axis.title.x = element_text(size = 12, face = "bold"),
    axis.title.y = element_text(size = 12, face = "bold"),
    axis.text.x = element_text(size = 8, face = "bold"),
    axis.text.y = element_text(size = 8, face = "bold"),
    legend.text = element_text(size = 12, face = "bold"),
    legend.title = element_text(size = 13, face = "bold"),
    legend.key.size = unit(1.5, "lines"),
    legend.position = "right",
    strip.text = element_text(size = 12, face = "bold")) +
  xlab("Variation Value") +
  ylab("Parameter")

H_clust$top50_parameter


```


```{r}


############################################################################################################################################
#                                                                                                                                          #
#                                                   Plotting parameter values across different cluster                                     #
#                                                                                                                                          #
############################################################################################################################################

# In some cases you need to adjust the y-limit to get a proper distribution of the graph 

ggplot(PCA$df_pca, aes(x = as.factor(Cluster), y = RI, group = Cluster, fill = as.factor(Cluster))) +
  # add half-violin from {ggdist} package
  stat_halfeye(
    # adjust bandwidth
    adjust = 1,
    # move to the right
    justification = -0.05,
    # remove the slub interval
    .width = 0,
    point_colour = NA
  ) +
  geom_boxplot(
    width = 0.10,
    # removing outliers
    outlier.color = NA,
    alpha = 0.3
  ) +
  scale_fill_manual(values = company_colors) +
  xlab("") +
  theme_bw() +
  #ylim(c(0,3000))+
  labs(fill = "Cluster") +
  theme(
    plot.title = element_text(size = 14, hjust = 0.5, face = "bold"),
    axis.title.x = element_text(size = 12, face = "bold"),
    axis.title.y = element_text(size = 12, face = "bold"),
    axis.text.x = element_text(size = 8, face = "bold"),
    axis.text.y = element_text(size = 8, face = "bold"),
    legend.text = element_text(size = 12, face = "bold"),
    legend.title = element_text(size = 13, face = "bold"),
    legend.key.size = unit(1.5, "lines"),
    legend.position = "right",
    strip.text = element_text(size = 12, face = "bold"))
  



```


```{r}

############################################################################################################################################
#                                                                                                                                          #
#                                                   Plotting proportion of differing morphologies                                          #
#                                                                                                                                          #
############################################################################################################################################

##### Overall trend for morpho-frequence heatmap #####
Morpho <- list()

Morpho$df_Morpho_count <- PCA$df_pca[,c(4, 5, 6,14, 62)]
Morpho$df_Morpho_count.t <- table(Morpho$df_Morpho_count)
Morpho$df_Morpho_count_scale.t <- data.frame(Morpho$df_Morpho_count.t)


##### new merge of clutsrer

Morpho$df_Morpho_count$Cluster_new <- case_when(Morpho$df_Morpho_count$Cluster=="1" ~ "Large Ameboid",
                                                Morpho$df_Morpho_count$Cluster=="2" ~ "Transition",
                                                Morpho$df_Morpho_count$Cluster=="3" ~ "Large Ameboid",
                                                Morpho$df_Morpho_count$Cluster=="4" ~ "Homeostatic",
                                                Morpho$df_Morpho_count$Cluster=="5" ~ "Homeostatic",
                                                Morpho$df_Morpho_count$Cluster=="6" ~ "Homeostatic")

##### Proportion Maps for major 4 phenotypes #####

Morpho$df_morpho_weeks <- Morpho$df_Morpho_count

# Calculate the counts and proportions for each bin number and phenotype
Morpho$proportion_count <- table(Morpho$df_morpho_weeks)

#define function to scale values between 0 and 1
Morpho$scale_values <- function(x){(x-min(x))/(max(x)-min(x))}

Morpho$proportion_count <- data.frame(Morpho$scale_values(Morpho$proportion_count))


# Plot the proportion bar plot for overall comparison
g1 <- ggplot(Morpho$proportion_count, aes(fill = Cluster_new, y = Freq, x = MW)) + 
  geom_bar(position = "fill", stat = "identity") +
  facet_grid(Expo_Time ~ Recovery_Time) +
  xlab("") +
  ylim(c(0,1.1)) +
  ylab("Proportion of Cells") +
  labs(fill = "Phenotype") +
  scale_fill_manual(values = company_colors) +
  ggtitle(" Morphology proportion overall") +
  theme_classic() +
  ggpubr::stat_compare_means(comparisons = list(c("Sham","Stimulation")), method = "wilcox.test") +
  theme(
    plot.title = element_text(size = 12, hjust = 0.5, face = "bold"),
    axis.title.x = element_text(size = 12, face = "bold"),
    axis.title.y = element_text(size = 12, face = "bold"),
    axis.text.x = element_text(size = 10, face = "bold", angle = 45, hjust = 1),
    axis.text.y = element_text(size = 10, face = "bold"),
    legend.text = element_text(size = 9, face = "bold"),
    legend.title = element_text(size = 9, face = "bold"),
    legend.key.size = unit(1.5, "lines"),
    legend.position = "NA",
    strip.text = element_text(size = 10, face = "bold")
  )

# Plot the proportion bar plot for center vs periphery comparison
g2 <- ggplot(Morpho$proportion_count, aes(fill = Cluster_new, y = Freq, x = MW)) + 
  geom_bar(position = "fill", stat = "identity") +
  facet_grid(Region ~ Recovery_Time) +
  xlab("") +
  ylim(c(0,1.1)) +
  ylab("Proportion of Cells") +
  labs(fill = "Phenotype") +
  scale_fill_manual(values = company_colors) +
  ggtitle("Morphology proportion center vs periphery") +
  ggpubr::stat_compare_means(comparisons = list(c("Sham","Stimulation")), method = "wilcox.test") +
  theme_classic() +
  theme(
    plot.title = element_text(size = 12, hjust = 0.5, face = "bold"),
    axis.title.x = element_text(size = 12, face = "bold"),
    axis.title.y = element_text(size = 12, face = "bold"),
    axis.text.x = element_text(size = 10, face = "bold", angle = 45, hjust = 1),
    axis.text.y = element_text(size = 10, face = "bold"),
    legend.text = element_text(size = 9, face = "bold"),
    legend.title = element_text(size = 9, face = "bold"),
    legend.key.size = unit(1.5, "lines"),
    legend.position = "right",
    strip.text = element_text(size = 10, face = "bold")
  )

g1+g2

```

```{r}


############################################################################################################################################
#                                                                                                                                          #
#                                                                Correlation heatmap                                                       #
#                                                                                                                                          #
############################################################################################################################################


# Correlation Plot

Corr <- list()

# Compute the correlation matrix
Corr$Corr_df <- cor(PCA$df_pca[, 15:61])

# Define a function to convert distance to order
dist2order <- function(corr, method, ...) {
  d_corr <- as.dist(1 - corr)
  s <- seriate(d_corr, method = method, ...)
  i <- get_order(s)
  return(i)
}

# Multidimensional Scaling and ordering
i <- dist2order(Corr$Corr_df, 'HC_ward')

# Plot the correlation matrix
corrplot(Corr$Corr_df[i, i], cl.pos = 'n')


```




